---
title: 'Identifying immune features that determine immunogenicity and influence prognosis in chemotherapy treated diffuse large B cell lymphoma (DLBCL)'
author: 'Shruti Sridhar'
date: 'May 2022'
institution: 'Reed College'
division: 'Mathematics and Natural Sciences'
advisor: 'Advisor F. Name'
# If you have more two advisors, un-silence line 7
#altadvisor: 'Your Other Advisor'
department: 'Biological Sciences'
degree: 'Bachelor of Arts'
knit: bookdown::render_book
site: bookdown::bookdown_site

# The next two lines allow you to change the spacing in your thesis. You can 
# switch out \onehalfspacing with \singlespacing or \doublespacing, if desired.
header-includes:
    - \usepackage{setspace}\onehalfspacing

# This will automatically install the {remotes} package and {thesisdown}
# Change this to FALSE if you'd like to install them manually on your own.
params:
  'Install needed packages for {thesisdown}': True
  
# Remove the hashtag to specify which version of output you would like.
# Can only choose one at a time.
output:
# thesisdown::thesis_pdf: default 
  thesisdown::thesis_gitbook: default         
#  thesisdown::thesis_word: default
#  thesisdown::thesis_epub: default

# If you are creating a PDF you'll need to write your preliminary content 
# (e.g., abstract, acknowledgements) below or use code similar to line 25-26 
# for the .RMD files. If you are NOT producing a PDF, delete or silence
# lines 25-39 in this YAML header.
abstract: '`r if(knitr:::is_latex_output()) paste(readLines(here::here("prelims", "00-abstract.Rmd")), collapse = "\n  ")`'
# If you'd rather include the preliminary content in files instead of inline
# like below, use a command like that for the abstract above.  Note that a tab 
# is needed on the line after the `|`.
acknowledgements: |
  I want to thank a few people.
#dedication: |
  #You can have a dedication here if you wish. 
#preface: |
  #This is an example of a thesis setup to use the reed thesis document class 
  #(for LaTeX) and the R bookdown package, in general.
  
# Specify the location of the bibliography below
bibliography: bib/thesis.bib
# Download your specific csl file and refer to it in the line below.
csl: csl/apa.csl
lot: true
lof: true
---

<!--
Above is the YAML (YAML Ain't Markup Language) header that includes a lot of 
metadata used to produce the document.  Be careful with spacing in this header!

If you'd prefer to not include a Dedication, for example, simply delete the section entirely, or silence them (add # before each line). 

If you have other LaTeX packages you would like to include, delete the # before header-includes and list the packages after hyphens on new lines.

If you'd like to include a comment that won't be produced in your resulting file enclose it in a block like this.

If you receive a duplicate label error after knitting, make sure to delete the index.Rmd file and then knit again.
-->

```{r include_packages, include=FALSE}
# This chunk ensures that the thesisdown package is
# installed and loaded. This thesisdown package includes
# the template files for the thesis.
if (!require(remotes)) {
  if (params$`Install needed packages for {thesisdown}`) {
    install.packages("remotes", repos = "https://cran.rstudio.com")
  } else {
    stop(
      paste('You need to run install.packages("remotes")",
            "first in the Console.')
    )
  }
}
if (!require(thesisdown)) {
  if (params$`Install needed packages for {thesisdown}`) {
    remotes::install_github("ismayc/thesisdown")
  } else {
    stop(
      paste(
        "You need to run",
        'remotes::install_github("ismayc/thesisdown")',
        "first in the Console."
      )
    )
  }
}
library(thesisdown)
# Set how wide the R output will go
options(width = 70)
```

<!--
The acknowledgments, preface, dedication, and abstract are added into the PDF
version automatically by inputting them in the YAML at the top of this file.
Alternatively, you can put that content in files like 00--prelim.Rmd and
00-abstract.Rmd like done below.
-->



```{r eval=!knitr::is_latex_output(), child=here::here("prelims", "00--prelim.Rmd")}

```

```{r eval=!knitr::is_latex_output(), child=here::here("prelims", "00-abstract.Rmd")}

```

<!-- The {.unnumbered} option here means that the introduction will be 
"Chapter 0." You can also use {-} for no numbers on chapters.
-->

# Introduction {.unnumbered}

Modulation of the immune system plays a key role in modern cancer treatment¹. DLBCL is the most common aggressive lymphoma, affecting both young and elderly adults². The prognosis of DLBCL is influenced by the immune microenvironment which plays a key role in the prognosis of DLBCL³. For instance, M2 macrophages might have a lymphoma-promoting function in DLBCL and predict poor clinical outcomes, unlike M1 macrophages⁴ .Using bioinformatic techniques we can then gain insight into the pathways that regulate the infiltration of various immune cells.' 
 
To facilitate the investigation of the roles of various immune cells in tumors, various computational methods5,6,7,8,9,10,11,12 have been developed, particularly for deconvoluting bulk gene expression profiles that will allow us to estimate various cell compositions. The idea behind such computational approaches is to treat bulk expression profiles as a linear or additive combination of the expression profiles of several cell subsets, and thus recover the fractions of different types of cells. Applying immune deconvolution algorithms on DLBCL datasets would allow us to examine the fraction of immune cells present in tumor samples and hence help us gauge the extent of immune cell infiltration. Further downstream bioinformatic analysis such as Gene set enrichment or pathway analysis can help us gain insight into the pathways that regulate the infiltration of various immune cells. 
 
 
Chemotherapy treatment has also been found to play a role in immune cell recruitment. A key mediator of chemotherapy-induced immune signaling is the Stimulator of interferon genes (STING). The cGAS-STING pathway mediates the interplay between the cytotoxic effect and immune stimulation by detecting DNA damage-induced micronuclei or cytoplasmic chromatin fragments and promoting antitumor immune responses13. Understanding the role of STING in mediating an immune response as a cytoplasmic nucleic acid sensor would give us insight into the immune microenvironment in the context of DLBCL. Further, we have preliminary evidence that indicates that STING promote a favorable	 prognosis in DLBCL thus we reasoned that this protein could be a key determinant of immunogenicity in DLBCL

